% !TEX encoding = UTF-8 Unicode
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
%%
%%  This file is asmeconf-template.tex, a LaTeX template to format ASME Conference papers according to
%%  the requirements on ASME's conference web pagess, and including hypertext support for the pdf.
%%
%%  This file is version 1.45 dated 2025/11/10
%%  
%%  As of version 1.11, this template defaults to ASME's newer conference guidelines first posted July 2019.
%% 			Those guidelines changed the author block formatting to be inline. 
%%			If you prefer the traditional grid format, use the class option [grid].
%%			Nomenclature now follows the abstract, and the abstract text is set in italics.
%%
%%  Author: John H. Lienhard, V
%%          Department of Mechanical Engineering
%%          Massachusetts Institute of Technology
%%          Cambridge, MA 02139-4307 USA
%%
%%  Class options include:
%%
%%          * An option to use the traditional grid arrangement of author names, [grid].  
%%			*	 With this option, line breaks (\\) may be inserted into the address as needed. 
%%			*	 Author names that include commas should be enclosed in braces, e.g.,  {Joseph L. Smith, Jr.}.  
%%			*	 Authors may be grouped above a single affiliation using braces, e.g., {Henry Tudor, Catherine Parr}.
%%
%%          * An option to balance the heights of columns on the last page, [balance]. 
%%
%%          * An option to include line numbers, [lineno]. You must *run twice* for proper placement of the line numbers. 
%%			*	 The lineno package does not number titles, footnotes, captions, or tables.
%%          *    This option will disable balancing column height on final page if that option has been invoked.
%%          *    The lineno package won't always number the lines preceding displayed math in a paragraph because
%%          *    paragraph has not ended.  See that package's documentation for macros to address this problem, or
%%          *    just leave a blank line above the displayed equation while you are editing and then remove the 
%%          *    blank line along with [lineno] option when you move to your final version.
%%
%%			* An option not to use a boldface font for caption text, [unboldcaption]
%%
%%          * Options to: 	omit the ASME copyright footer, [nofoot]; 
%%			*				use government employee copyright notice,      [govt];
%%			*				use government contractor copyright notice,    [contractor];
%%			*				use copyright notice for some gov't employees, [somegovt];
%%			*				omit the conference headers on the titlepage,  [nohead]
%%
%%			* Option to use LuaLaTeX **without** unicode-math and fontspec, [nofontspec]
%%
%%          * Math options from M. Sharpe's newtxmath package: upright integrals [upint]; 
%%          *    [varvw] for a v and w that are better distinguished from Greek nu; and also 
%%          *    [subscriptcorrection, smallerops, varg, frenchmath, varbb, cmbraces, slantedGreek,...] 
%%			* 	 See newtx documentation for descriptions (at CTAN: https://www.ctan.org/pkg/newtx, v1.744 or higher is best).
%%			*	 newtx is not loaded with lualatex unless the [nofontspec] option is used.
%%
%%          * An option to allow hyphenation of the typewriter font [hyphenate], from the inconsolata package (pdfTeX only).
%%          *    Hyphenation is normally suppressed for a typewriter (monospaced) font because those are often used for code.
%%			*	 To replace the default variable word spacing by monospacing, use the option [mono].
%%			*	 To get a zero without a slash, use [var0]
%%
%%			* PDF/A compliance:  Since 2022, LaTeX has included support for PDF/A, through the \DocumentMetadata{..} command.
%%
%%          * Options (used by the babel package) to include passages in languages other than English (e.g., a translation 
%%			*    of the abstract).  See Appendix C for details.
%%			*    	pdftex: Appropriate scripts will be loaded if you call the class options [greek], [russian], or [vietnamese].
%%			*    
%%			*    	lualatex: Language support is most extensive when running LuaLaTeX, which automatically loads fontspec.
%%			*		Non-Latin scripts require the [loadscripts] option, see Appendix C. 
%%			*	 	Be aware that you might need to install additional fonts for some languages.
%%			 
%%  The use of commands defined or modified by the asmeconf class is illustrated throughout this file. In particular, 
%%  ASME requires capitalized section headings, and as a result some care is needed when using macros in section headings,
%%  as also illustrated below.
%%
%%  Use an **up-to-date and complete** installation of LaTeX, such as TeX Live 2025 or later. 
%%		LaTeX formats earlier than 2022 are not fully supported.
%%
 %=========================================================
%% 
%% LICENSE: 
%%
%% Copyright (c) 2025 John H. Lienhard
%%
%% Offered under the MIT license: https://ctan.org/license/mit 
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 

%%  RECOMMENDED pdf management code.
%%  This code was added to the LaTeX kernel in June 2022.
%% 		See https://www.latex-project.org/news/latex2e-news/ltnews35.pdf
%%  If you have problems with these lines, you can comment them out or, better, update your LaTeX format.

\DocumentMetadata{
	lang		= eng-US, 	% default
	pdfstandard	= A-3u,		% A-2b, A-2u, A-3b, A-3u. Check that your figures also meet the standard you choose.
	pdfversion	= 1.7, 		% an older version of PDF, but very widely supported
%	pdfversion  = 2.0, 		% default for LaTeX. Version 2.0 supports tagged PDF.
%	pdfstandard = { ua-2 , a-4f }, % run with lualatex and use a very recent LaTeX format (e.g., 2025/11/01)
%	tagging 	= on, 		% for producing tagged pdf
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% CLASS OPTIONS are described above. Change the options given below to meet your needs.
%% 
%%	 	NB: Remove the [colorlinks] option before *final* submission to ASME, to get black text for printing,
%%			but keep that option for electronic use.
%%
%%		NB: If you are not using the language options, * remove * them (together with Appendices C and D).
%%			Greek, cyrillic languages, and vietnamese if used, must be named as \documentclass options under pdftex.
%%			Spanish, german, and many others, if used, do not need to be named when the babel package is dated 2024 or later.
%%
%%		NB: the mathalfa option was dropped in v1.41; load the mathalpha package in your preamble instead.

\documentclass[colorlinks,upint,subscriptcorrection,varvw,hyphenate,balance]{asmeconf} % pdftex
\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning,calc,shapes.geometric}
\usepackage{placeins}
\usepackage{float}

%\documentclass[colorlinks,upint,subscriptcorrection,varvw,balance,loadscripts,german,vietamese]{asmeconf} % lualatex

%%%%%  pdf metadata  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\hypersetup{%
	pdfauthor={John H. Lienhard},									  % <=== change to YOUR name
	pdftitle={ASME Conference Paper LaTeX Template},                  % <=== change to YOUR pdf file title
	pdfkeywords={ASME conference paper, LaTeX template, BibTeX style},% <=== change to YOUR pdf keywords
	pdfsubject = {Describes the asmeconf LaTeX template},			  % <=== change to YOUR subject
%	pdfurl={https://ctan.org/pkg/asmeconf},% may delete
%	pdflicenseurl={https://ctan.org/pkg/asmeconf},% may delete
}
% If an author name or the title include a comma, enclosed it in braces, e.g., pdfauthor{John Forbes Nash{,} Jr.}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\allowdisplaybreaks % from amsmath package, allows multiline equations to break across pages (delete if not wanted)
					% using \\* instead of \\ will prevent specific lines from being pagebreaks.

% logos used in this document
\newcommand*\pdfTeX{pdf\TeX}    
\newcommand*\LuaLaTeX{Lua\LaTeX}
\newcommand*\BibTeX{Bib\TeX}

\begin{document}

% Change these fields to the right content for your conference.
% You can comment these out if for some reason you don't want a header.
% Use title case for the conference name (first letters capitalized), not all capitals

\ConfName{Proceedings of the ASME 2026\linebreak International Design Engineering Technical Conferences \&\linebreak Computers and Information in Engineering Conference}
\ConfAcronym{IDETC-CIE 2026}
\ConfDate{August 23--26, 2026}
\ConfCity{Houston, TX, USA}
\PaperNo{DETC2026-XXXX}

% Units of measure (e.g., cm) and other specialty lowercase terms in the title should be 
%   enclosed in \NoCaseChange{...} to maintain lower case type.
%   The rest of the title will automatically be set in all capital letters.
%
%	\title{Place Title Here: Place Subtitle After Colon} 

\title{Reliable Natural-Language to SysMLv2 Translation via Compiler-Driven Iterative Refinement} % <=== replace with YOUR title
 
%   Put author names into the order you want. Use the same order for affiliations.
%   \affil{#} tags the author's affiliation to the address in \SetAffiliation{#}.
%   No space between last name and \affil{#}, separate names with commas.
%
%	For a sole author or a single affiliation for all authors, {#} may be left empty, i.e. \affil{} and \SetAffiliation{} (but not with [grid] option!)
%
%   \CorrespondingAuthor{email} follows that author's affiliation, no spaces before or after 
%   If multiple corresponding authors, put both email addresses in the same command and place after both authors.
%
%   \JointFirstAuthor, if applicable, follows the affiliation of the relevant authors, no spaces.

\SetAuthors{%
    Chance LaVoie\affil{1}\CorrespondingAuthor{chancel@cmu.edu},
    Eladio Andujar Lugo\affil{1},
    Levent Burak Kara\affil{1}
}

\SetAffiliation{1}{Department of Mechanical Engineering, Carnegie Mellon University, Pittsburgh, PA 15213 USA}
%   Note: You can force a line break in the address using \\ 

%	To switch from inline author names to gridded names, use the [grid] option.

\maketitle

%%% Use this footnote for tracking various versions of your draft. Change text to suit your own needs. 
%%% \date{..} calls the same command. 
\versionfootnote{Documentation for \texttt{asmeconf.cls}: Version~\versionno, \today.}% <=== Delete before final submission.

%%% Change the following to your keywords.  Keywords are automatically printed at the end of the abstract.
%%% This command MUST COME BEFORE the end of the abstract.
%%% If you don't want keywords, leave the argument of \keywords{} empty (or use the abstract* environment)

\keywords{SysMLv2, model-based systems engineering, large language models, compiler-in-the-loop, natural language to model translation, syntactic validity}

%%%%%  End of fields to be completed. Now write your paper. %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


%%%%%  ABSTRACT  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% Abstract should be 200 words or less
\begin{abstract}
SysMLv2 is a textual, compilable modeling language, so generated output is only useful if it passes deterministic compiler checks. One-shot large language model generation frequently produces plausible but uncompilable SysMLv2, motivating compiler-gated refinement rather than one-pass synthesis.

This paper studies a syntax-only compiler-in-the-loop pipeline on the full SysMBench prompt set (IDs 1--151), evaluated across four model configurations. For each prompt/model trial, we compare baseline single-shot output (iteration 1 only) against the final output after iterative generate--compile--repair using \texttt{syside check} as the oracle. Across 604 trials, first-shot compilation succeeds on 51.16\% (309/604; 95\% Wilson CI: 47.18--55.13\%), while the compiler-in-the-loop final output reaches 100.00\% eventual compilation (604/604; 95\% Wilson CI: 99.37--100.00\%), a gain of 48.84 percentage points with zero unresolved prompts. Failed-first-shot cases recover fully (295/295). These results show that deterministic compiler feedback is the primary control signal for reliable natural-language--to--SysMLv2 generation and that single-shot generation leaves a large, recoverable syntactic error surface.
\end{abstract}

%%%%%%%%%  BODY OF PAPER %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}
Model-Based Systems Engineering (MBSE) relies on formal, machine-checkable artifacts to support early design iteration, requirements traceability, and verification. The release of Systems Modeling Language version~2 (SysMLv2)~\cite{omgSysMLv2Spec2024} strengthens this workflow...
  strengthens this workflow by introducing a fully textual modeling language with a well-defined grammar and an executable toolchain, enabling models to be parsed, compiled, and version-controlled like software.

This shift creates an opportunity for automated model synthesis: engineers could describe a system in natural language and obtain a SysMLv2 model that can be compiled and analyzed. Large language models (LLMs) are a natural candidate for this translation task, given their success at generating structured code from natural-language specifications. However, a model that is ``almost correct'' is not useful in practice: even small syntax or metamodel errors prevent compilation and therefore block rendering, visualization, and downstream analysis. Reliable translation therefore requires \emph{syntactic validity}, not just plausible-looking output.

The current status quo is that LLMs can often produce readable pseudo-code or SysML-like text. Pseudo-code is not sufficient for MBSE: SysML is intended to serve as a \emph{compilable} and \emph{visualizable} system representation, and compilation is the gateway to rendering, visualization, and analysis inside engineering toolchains. If an LLM output is not syntactically correct, it cannot be compiled and therefore cannot be rendered into the diagrams and model views that engineers use to understand and communicate system structure and behavior. As a result, the practical value of AI-assisted modeling hinges on producing syntactically valid SysMLv2, not informal approximations.

In many code-generation domains, syntactic reliability can be improved by training or fine-tuning on large corpora of valid programs. For SysMLv2, such data is currently scarce: there is no large, broadly available dataset of compiler-verified models, and the language itself is evolving. Consequently, one-shot LLM generations often mix SysML~v1 patterns with SysMLv2 constructs or violate the SysMLv2 grammar, leading to outputs that read convincingly but fail deterministic checks. This data scarcity motivates a method that improves correctness \emph{without} modifying model weights.

This paper presents a compiler-in-the-loop framework that couples an off-the-shelf LLM with a deterministic verifier in an iterative generate--compile--repair loop. At each iteration, the model proposes a complete SysMLv2 candidate; the verifier returns precise diagnostics (error types and locations); and a controller converts these diagnostics into targeted revision instructions. The loop terminates only when zero errors are reported, yielding syntactic validity by construction without training, fine-tuning, or specialized hardware.

We evaluate the approach on all 151 SysMBench prompts and compare two conditions for each prompt/model pair: single-shot compilation at iteration 1 versus final compilation after compiler-in-the-loop refinement. Results show that deterministic compiler feedback is the primary driver of dependable generation: iterative refinement converts many first-pass failures into compilable SysMLv2 while single-shot generation alone leaves a large recoverable error surface.

\section{Related Work}

Our review focuses on prior efforts in automated SysML v2 generation, grammar-constrained structured synthesis, and compiler-guided iterative refinement, with an emphasis on approaches addressing reliable syntactic correctness in low-data modeling languages.

\subsection{LLM-Based SysML v2 Model Generation}
Recent work has explored using large language models to generate SysML v2 models from natural language. Bouamra et al.~\cite{bouamra2025systemp} propose SysTemp, a multi-agent framework that decomposes the generation task into requirement extraction, template-based skeleton construction, and grammar-level parsing feedback. Their system employs a ParserAgent to detect syntactic errors and iteratively repair generated models, achieving 80\% convergence across five evaluation scenarios. This work demonstrates that structural scaffolding and agent decomposition improve grammar conformity in sparse-data settings. However, syntactic correctness is defined at the grammar level and convergence is reported empirically rather than enforced as a termination invariant.

Cibrián et al.~\cite{cibrian2025sysmlagent} introduce an agent-based framework combining retrieval-augmented generation (RAG) with ANTLR-based grammar validation for SysML v2 synthesis. Their approach reports 100\% syntactic validity across 20 curated prompts under grammar-level parsing. While this represents a significant advance in structured generation reliability, correctness is enforced through grammar parsing rather than full compiler validation. Because grammar-level validation ensures context-free structural conformity but does not enforce full static semantics under the official toolchain, grammar-valid models may still fail compilation in production environments.
neccessary but not sufficient
Together, these efforts establish that iterative validation substantially improves syntactic success in SysML v2 generation. However, prior systems define correctness at the grammar level and evaluate on limited scenario sets. Our work builds on these insights while shifting the correctness oracle from grammar parsing to the official SysIDE compiler and scaling evaluation to benchmark-level scenarios.

\subsection{Grammar-Constrained and Template-Based Structured Synthesis}

Structured synthesis approaches aim to reduce hallucinations in LLM output by constraining generation via templates or grammar rules. SysTemp~\cite{bouamra2025systemp} explicitly uses a template generator based on Jinja2 to construct syntactically compliant model skeletons prior to completion. Grammar-constrained decoding and post-generation parsing similarly reduce token-level structural invalidity.

While grammar validation ensures adherence to context-free rules, it does not enforce type resolution, cross-reference integrity, constraint satisfaction, or toolchain compatibility. Thus, grammar-valid artifacts may remain unusable within industrial MBSE workflows. Our approach differs by treating grammar conformity as necessary but insufficient and requiring full compiler acceptance prior to termination.

\subsection{Compiler-in-the-Loop and Verifier-Guided Generation}

Prior work in neural code generation has explored leveraging compiler diagnostics to improve the compilability of model outputs. For example, Wang et al.~\cite{wang2022compilable} propose a multi-stage refinement framework that uses compiler feedback to iteratively revise generated programs and increase compilation success rates. Such approaches demonstrate that deterministic compiler signals can serve as effective supervisory feedback in programming-language settings. However, these systems operate in mature programming ecosystems and treat compilation success as an empirical objective rather than as a structural termination invariant.

Beyond compiler-feedback approaches in code generation, iterative generation guided by deterministic verifiers has also been studied under counterexample-guided inductive synthesis and execution-based refinement paradigms. Grubišić et al.~\cite{grubisic2024compiler} demonstrate that LLVM compiler feedback can serve as a supervisory signal for refining LLM-generated intermediate representation. Their work shows that deterministic compilation signals constrain output space and improve structural validity in programming languages.

While these approaches establish the feasibility of compiler-aware refinement in software domains, they differ in both objective and context from model-based systems engineering. Compiler feedback in such systems is typically used to improve optimization quality or increase compilation probability, rather than to enforce strict termination conditions. Moreover, these methods operate in programming languages with extensive training corpora and stable ecosystems.

In contrast, our work applies compiler-in-the-loop refinement to SysMLv2, a newly standardized modeling language with sparse representation in LLM training data and strict toolchain requirements. Rather than treating compilation results as heuristic feedback, we enforce zero-error compilation under the SysMLv2 compiler (invoked via \texttt{syside check}) as a termination invariant. This elevates syntactic correctness from an empirical metric to a property guaranteed by construction, aligning generation reliability with production MBSE toolchain criteria rather than grammar-level approximations.


\input{methods/methods_syntactic_only}

\input{results/results_syntactic_only}

\section{Discussion}
We discuss why deterministic compiler diagnostics provide an effective supervision signal for new, sparsely represented languages like SysMLv2, and the resulting trade-offs in runtime and number of iterations.

\section{Conclusion}
We present a reproducible pathway for reliable natural-language to SysMLv2 translation by embedding deterministic compiler feedback into LLM generation, thereby guaranteeing syntactically valid (compilable) output by construction.

\section*{Acknowledgments}

\bibliographystyle{asmeconf}
\bibliography{references}

\end{document}
